MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-05-12 latex/paper/sections/05-7series_architecture.tex
more packing

Figure~\ref{fig:hierarchy_4} illustrates a high-level view of the hierarchical organization of a 7-Series FPGA. 
At its lowest level, the device consists of a large array of atomic components called \emph{Basic Elements of Logic}~(\textbf{BELs}). 
These BELs encompass look-up tables (\textbf{LUTs}), flip-flops (\textbf{FFs}), block RAMs (\textbf{BRAMs}), DSP slices (\textbf{DSPs}), and the configurable interconnect fabric. 
They constitute the fundamental building blocks for implementing digital circuits on the FPGA.
Note how the Tile arrangement is columnar such that column consists of only one Tile type.

To manage this complexity, Xilinx organizes these BELs into incrementally abstract structures. 
First, \textbf{BELs} are grouped into \textbf{Sites}. 
Each Site is embedded into a \textbf{Tile}, and Tiles are further arranged into \textbf{Clock Regions}. 
In some high-density devices, multiple Clock Regions may be consolidated into one or more \textbf{Super Logic Regions}~(SLRs). 
However, for the scope of this paper, we focus on Xilinx 7-Series devices with only a single SLR.

@@

In the 7-Series architecture, the term \emph{CLB} (Configurable Logic Block) refers to a \emph{CLB Tile} that contains two \emph{SLICE} Sites. 
Xilinx offers two variants of SLICE Sites: \textbf{SLICEL} and \textbf{SLICEM}.
\begin{itemize}
\item Each \texttt{SLICEL} has a set of BELs including eight \texttt{LUT}s, eight \texttt{FF}s, and one \texttt{CARRY4} adder. The \texttt{LUT} BELs in a \texttt{SLICEL} can only host \texttt{LUT} Cells. 
\item The \texttt{SLICEM} includes all the features of a \texttt{SLICEL} but its \texttt{LUT} BELs can host both \texttt{LUT} Cells, which are asynchronous ROM elements, or \texttt{RAM32M} Cells, which are synchronous 32-deep RAM elements. These cells are also referred to as \emph{Distributed RAM} in the Xilinx documentation.
    These cells can offer an alternative to the larger, more dedicated 18K-36K \texttt{RAMB18E1} cells when RAM resources are highly utilized. 
\end{itemize}
In a typical 7-Series device, approximately 75\% of the SLICE Sites are \texttt{SLICEL}s and 25\% are \texttt{SLICEM}s. 
A single CLB Tile can therefore host either two \texttt{SLICEL}s or one \texttt{SLICEL} and one \texttt{SLICEM}.
To simplify the problem space, however, we will only consider \texttt{SLICEL}s for general logic and use the dedicated \texttt{RAMB18E1} cells to implement RAM elements.

@@

{
    \centering
    \includegraphics[width=\columnwidth]{figures/carry_cell_group.png}
    \captionof{figure}{A \texttt{SLICEL} with a \texttt{CARRY4} cell, 4 LUT cells, and 4 FF cells placed inside as shown in the Vivado device viewer}
    \label{fig:carry_cell_group}
}

{
    \centering
    \includegraphics[valign=c, width=4.0cm]{figures/carry_chain_placement.png}
    \includegraphics[valign=c, width=4.0cm]{figures/carry_chain_routes_3.png}
    \captionof{figure}{
        A \texttt{CARRY4} chain of size 3 placed across 3 SLICEs.
        \textbf{Left:} Simplified view, \textbf{Right:} As shown in the Vivado device viewer.
    }
    \label{fig:carry_cell_group}
}

>>>>

Figure~\ref{fig:hierarchy_4} illustrates a high-level view of the hierarchical organization of a 7-Series FPGA. 
At its lowest level, the device consists of a large array of atomic components called \emph{Basic Elements of Logic}~(\textbf{BELs}). 
These BELs encompass look-up tables (\textbf{LUTs}), flip-flops (\textbf{FFs}), block RAMs (\textbf{BRAMs}), DSP slices (\textbf{DSPs}), and the configurable interconnect fabric. 
They constitute the fundamental building blocks for implementing digital circuits on the FPGA.

To manage this complexity, Xilinx organizes these BELs into incrementally abstract structures. 
First, \textbf{BELs} are grouped into \textbf{Sites}. 
Each Site is embedded into a \textbf{Tile}, and Tiles are further arranged into \textbf{Clock Regions}. 
Note how the Tile arrangement is columnar such that column consists of only one Tile type. 
In some high-density devices, multiple Clock Regions may be consolidated into one or more \textbf{Super Logic Regions}~(SLRs). 
However, for the scope of this paper, we focus on Xilinx 7-Series devices with only a single SLR.

@@

In the 7-Series architecture, the term \emph{CLB} (Configurable Logic Block) refers to a \emph{CLB Tile} that contains two \emph{SLICE} Sites. 
Xilinx offers two variants of SLICE Sites: \textbf{SLICEL} and \textbf{SLICEM}.
\begin{itemize}
\item Each \texttt{SLICEL} has a set of BELs including eight \texttt{LUT}s, eight \texttt{FF}s, and one \texttt{CARRY4} adder. The \texttt{LUT} BELs in a \texttt{SLICEL} can only host \texttt{LUT} Cells. 
\item The \texttt{SLICEM} includes all the features of a \texttt{SLICEL} but its \texttt{LUT} BELs can host both \texttt{LUT} Cells, which are asynchronous ROM elements, or \texttt{RAM32M} Cells, which are synchronous 32-deep RAM elements. These cells are also referred to as \emph{Distributed RAM} in the Xilinx documentation.
    These cells can offer an alternative to the larger, more dedicated 18K-36K \texttt{RAMB18E1} cells when RAM resources are highly utilized or when the design inheritly demands homogeneously distributed local memory. 
\end{itemize}
In a typical 7-Series device, approximately 75\% of the SLICE Sites are \texttt{SLICEL}s and 25\% are \texttt{SLICEM}s. 
A single CLB Tile can therefore host either two \texttt{SLICEL}s or one \texttt{SLICEL} and one \texttt{SLICEM}.
To simplify the problem space, however, we will only consider \texttt{SLICEL}s for general logic and use the dedicated \texttt{RAMB18E1} cells to implement RAM elements.

@@

{
    \centering
    \includegraphics[width=\columnwidth]{figures/carry_cell_group.png}
    \captionof{figure}{A \texttt{SLICEL} populated by a \texttt{CARRY4} cell, 4 LUT cells, and 4 FF cells}
    \label{fig:carry_cell_group}
}

\newcolumn
{
    \centering
    \includegraphics[valign=c, width=4.0cm]{figures/carry_chain_placement.png}
    \includegraphics[valign=c, width=4.0cm]{figures/carry_chain_routes_3.png}
    \captionof{figure}{
        A \texttt{CARRY4} chain of size 3 placed across 3 SLICEs.
        \textbf{Left:} Simplified view, \textbf{Right:} As shown in the Vivado device viewer.
    }
    \label{fig:carry_cell_group}
}

### 2025-05-12 latex/paper/sections/08-netlist.tex
more packing

In either synthesized netlist, the \textbf{leaf cells}, (deepest level cells), must necessarily consist only of \textbf{primitive cells} from the architecture's primitive cell library (\texttt{LUT6}, \texttt{FDRE}, \texttt{CARRY4}, \texttt{DSP48E1}, etc.). 
The netlist can be compiled and exported as a purely structural low-level Verilog file, or an Electrinic Design Interchange Format (EDIF) file, both describing the netlist explicitly as a list of logical cells connected by a list of wires.

@@

\begin{lstlisting}[language=java, caption={Netlist extraction and traversal}, label={lst:netlist_extract}]
Design design = Design.readCheckpoint("synth.dcp")
EDIFNetlist netlist = design.getNetlist();

>>>>

In either synthesized netlist, the \textbf{leaf cells}, (deepest level cells), must necessarily consist only of \textbf{primitive cells} from the architecture's primitive cell library (\texttt{LUT6}, \texttt{FDRE}, \texttt{CARRY4}, \texttt{DSP48E1}, etc.). 
The netlist can be compiled and exported as a purely structural low-level Verilog file, or an Electrinic Design Interchange Format (EDIF) file, both describing the netlist explicitly as a list of logical cell instances connected by a list of wires.

@@

\begin{lstlisting}[language=java, caption={Basic netlist extraction and traversal}, label={lst:netlist_extract}]
Design design = Design.readCheckpoint("synth.dcp")
EDIFNetlist netlist = design.getNetlist();

### 2025-05-12 latex/paper/sections/09-packing.tex
more packing

// Select only the carry cells.
List<EDIFCellInst> carryCells = new ArrayList<>();
for (EDIFCellInst eci : ecis) {
    if (eci.getCellName().contains("CARRY4"))
        carryCells.add(eci);
}

@@

\subsection{Packing}
\label{subsec:packing}
Now that we have our \texttt{PrepackedDesign} object keeping track of multi-cell structures on the \texttt{edif} level, we can start packing them into \texttt{SiteInst} objects on the \texttt{design} level. 
Below are some of the most relevant classes from the \texttt{design} package for this task.
\begin{itemize}
    \item \texttt{Cell}: A cell corresponds to the leaf cell within the logical netlist \texttt{EDIFCellInst} and provides a mapping to a physical location BEL on the device. A cell can be created directly out of an \texttt{EDIFCellInst} to inherit all of its \texttt{edif} properties on the \texttt{design} level.
    \item \texttt{Net}: Represents the physical net to be routed (both inter-site and intra-site). When an \texttt{Cell} is created out of an \texttt{EDIFCellInst}, the \texttt{Net}s are automatically created out of its corresponding \texttt{EDIFNet}s.
    \item \texttt{SiteInst}: An instance of a \texttt{Site} on the \texttt{Device}. Carries the mapping information between the \texttt{BEL}s in a \texttt{Site} and the \texttt{Cell}s assigned to them. Also keeps track of the intra-Site routing information within (\texttt{Net}s, \texttt{SitePinInst}s, \texttt{SitePIP}s, etc.).
    \item \texttt{SitePIP}: Represents the fuses in routing \texttt{BEL}s. 
    \item \texttt{SitePinInst}: An instance of a \texttt{SitePin} on a \texttt{Site}. These objects serve as the interface between intra-Site routing and general inter-Site routing. 
\end{itemize}
\columnbreak

A \texttt{SiteInst} object can be created and its BELs populated by existing \texttt{EDIFHierCellInst} objects. 
A \texttt{SiteInst} is necessarily placed on a specific \texttt{Site} upon creation. 
Thus, the packer necessarily gives all \texttt{SiteInst}s an initial placement before randomization via simulated annealing in the following placement stage. 
In packing, will simply assign \texttt{SiteInst}s pseudorandomly onto device \texttt{Site}s coordinate order as they are created (e.g. first SiteInst onto \texttt{SLICEL\_X0Y0}, second \texttt{SiteInst} onto \texttt{SLICEL\_X0Y1}, etc.).

@@

Example:
    SiteInst si = new SiteInst("mySiteInst", design, SiteTypeEnum.SLICEL, device.getSite("SLICEL_X0Y1"));
    si.unplace();
    si.place(device.getSite("SLICEL_X15Y33"));
    // In Simulated Annealing, SiteInst objects will be unplaced() and placed() many times to converge to an optimal solution.

@@

\begin{lstlisting}[language=java, caption={Packing DSPCascades into \texttt{DSP48E1} \texttt{SiteInst}s.}, label={lst:carry_chains}]
private List<List<SiteInst>> packDSPCascades(List<List<EDIFHierCellInst>> EDIFDSPCascades) throws IOException {
    List<List<SiteInst>> siteInstCascades = new ArrayList<>();
    writer.write("\n\nPacking DSP Cascades... (" + EDIFDSPCascades.size() + ")");
    for (List<EDIFHierCellInst> cascade : EDIFDSPCascades) {
        List<SiteInst> siteInstCascade = new ArrayList<>();
        // each DSP tile has 2 DSP sites on the Zynq 7000
        writer.write("\n\tCascade Size: (" + cascade.size() + "), Cascade Anchor: "
                + cascade.get(0).getFullHierarchicalInstName());
        // Select a Site location on the device. 
        Site anchorSite = selectDSPAnchorSite(cascade.size());
        SiteTypeEnum siteType = SiteTypeEnum.DSP48E1;
        for (int i = 0; i < cascade.size(); i++) {
            writer.write("\n\t\tPlacing DSP: " + cascade.get(i).getFullHierarchicalInstName());
            Site site = (i == 0) ? anchorSite
                    : device.getSite("DSP48_X" + anchorSite.getInstanceX() + "Y" + (anchorSite.getInstanceY() + i));
            // Create the SiteInst object
            SiteInst si = new SiteInst(cascade.get(i).getFullHierarchicalInstName(), design, siteType, site);
            // Populate the SiteInst with a DSP Cell. 
            si.createCell(cascade.get(i), si.getBEL("DSP48E1"));
            occupiedSites.get(siteType).add(site);
            availableSites.get(siteType).remove(site);
            si.routeSite();
            siteInstCascade.add(si);
        }
        siteInstCascades.add(siteInstCascade);
    }
    return siteInstCascades;
} // end packDSPCascades()
\end{lstlisting}

\newpage
\begin{lstlisting}[language=java, caption={Packing LUTFFPairGroups into \texttt{SLICEL} \texttt{SiteInst}s}]
private List<SiteInst> packLUTFFPairGroups(
        Map<Pair<String, String>, LUTFFGroup> LUTFFEnableResetGroups) throws IOException {
    List<SiteInst> LUTFFSiteInsts = new ArrayList<>();
    writer.write("\n\nPacking LUT-FF Pair Groups... (" + LUTFFEnableResetGroups.size() + ")");
    for (Map.Entry<Pair<String, String>, LUTFFGroup> entry : LUTFFEnableResetGroups.entrySet()) {
        Pair<String, String> netPair = entry.getKey();
        String s1 = String.format("\n\tCENet: %-50s RNet: %-50s", netPair.key(), netPair.value());
        writer.write(s1);
        LUTFFGroup lutffgroup = entry.getValue();
        for (List<Pair<EDIFHierCellInst, EDIFHierCellInst>> LUTFFPairs : splitIntoGroups(lutffgroup.group(), 4)) {
            Site selectedSite = selectCLBSite();
            SiteInst si = new SiteInst(LUTFFPairs.get(0).value().getFullHierarchicalInstName(), design,
                    selectedSite.getSiteTypeEnum(), selectedSite);
            for (int i = 0; i < LUTFFPairs.size(); i++) {
                EDIFHierCellInst ff = LUTFFPairs.get(i).value();
                EDIFHierCellInst lut = LUTFFPairs.get(i).key();
                si.createCell(ff, si.getBEL(FF_BELS[i]));
                if (lut != null)
                    si.createCell(lut, si.getBEL(LUT6_BELS[i]));
            }
            si.routeSite();
            List<String> lutOPIP = new ArrayList<>(List.of("AUSED", "BUSED", "CUSED", "DUSED"));
            // DOES NOT SUPPORT STACKED LUTS OR LUT5s!
            for (int i = 0; i < LUTFFPairs.size(); i++) {
                EDIFHierCellInst lut = LUTFFPairs.get(i).key();
                if (lut != null) {
                    EDIFHierPortInst lutOPort = lut.getPortInst("O");
                    if (lutOPort.getHierarchicalNet().getLeafHierPortInsts(false, true).size() > 1) {
                        si.addSitePIP(si.getSitePIP(lutOPIP.get(i), "0"));
                    }
                }
            }
            rerouteFFClkSrCeNets(si);
            LUTFFSiteInsts.add(si);
        }
    }
    return LUTFFSiteInsts;
} // end packLUTFFPairGroups()
\end{lstlisting}

\begin{lstlisting}[language=java, caption={Packing LUTGroups into \texttt{SLICEL} \texttt{SiteInst}s}]
protected String[] LUT6_BELS = new String[] { "A6LUT", "B6LUT", "C6LUT", "D6LUT" };

private List<SiteInst> packLUTGroups(List<List<EDIFHierCellInst>> LUTGroups) throws IOException {
    List<SiteInst> LUTSiteInsts = new ArrayList<>();
    writer.write("\n\nPacking LUT Groups...");
    for (List<EDIFHierCellInst> group : LUTGroups) {
        Site selectedSite = selectCLBSite();
        SiteInst si = new SiteInst(group.get(0).getFullHierarchicalInstName(), design,
                selectedSite.getSiteTypeEnum(), selectedSite);
        for (int i = 0; i < group.size(); i++) {
            si.createCell(group.get(i), si.getBEL(LUT6_BELS[i]));
        }
        si.routeSite();
        // for whatever reason, this does not activate DUSED PIP for
        si.addSitePIP(si.getSitePIP("DUSED", "0"));
        si.addSitePIP(si.getSitePIP("CUSED", "0"));
        si.addSitePIP(si.getSitePIP("BUSED", "0"));
        si.addSitePIP(si.getSitePIP("AUSED", "0"));
        LUTSiteInsts.add(si);
    }
    return LUTSiteInsts;
} // end packLUTGroups()
\end{lstlisting}
\begin{multicols}{2}

\subsection{Placement}
    \label{subsec:placement}
    Up until now we have only organized the logical \texttt{EDIFCellInst}s into \texttt{SiteInst}s. 
    This is where simulated annealing actually begins where we actually place the \texttt{SiteInst}s onto physical \texttt{Site}s on the \texttt{device} level.

>>>>

// Select only the carry cells.
List<EDIFCellInst> carryCells = new ArrayList<>();
for (EDIFCellInst eci : ecis) {
    if (eci.getCellName().equals("CARRY4"))
        carryCells.add(eci);
}

@@

\subsection{Packing}
\label{subsec:packing}
Now that we have our \texttt{PrepackedDesign} object keeping track of multi-cell structures on the \texttt{edif} level, we can start packing them into \texttt{SiteInst} objects on the \texttt{design} level. 
Below are some of the most relevant classes from the \texttt{design} package for this task.
\begin{itemize}
    \item \texttt{Cell}: A cell corresponds to the leaf cell within the logical netlist \texttt{EDIFCellInst} and provides a mapping to a physical location BEL on the device. A cell can be created directly out of an \texttt{EDIFCellInst} to inherit all of its \texttt{edif} properties on the \texttt{design} level.
    \item \texttt{Net}: Represents the physical net to be routed (both inter-site and intra-site). When an \texttt{Cell} is created out of an \texttt{EDIFCellInst}, the \texttt{Net}s are automatically created out of its corresponding \texttt{EDIFNet}s.
    \item \texttt{SiteInst}: An instance of a \texttt{Site} on the \texttt{Device}. Carries the mapping information between the \texttt{BEL}s in a \texttt{Site} and the \texttt{Cell}s assigned to them. Also keeps track of the intra-Site routing information within (\texttt{Net}s, \texttt{SitePinInst}s, \texttt{SitePIP}s, etc.).
    \item \texttt{SitePIP}: A Programmable Interconnect Point (PIP) in a \texttt{Site}. Represents the fuses in intra-Site routing \texttt{BEL}s. 
    \item \texttt{SitePinInst}: An instance of a \texttt{SitePin} on a \texttt{Site}. These objects serve as the interface between intra-Site routing and general inter-Site routing. 
\end{itemize}
\columnbreak

A new \texttt{SiteInst} is created by populating its BELs with existing \texttt{EDIFHierCellInst} objects from the \texttt{EDIFNetlist}, and is immediately placed on a specific \texttt{Site}. 
The packer therefore assigns an initial placement to every \texttt{SiteInst} before the simulated‐annealing stage randomizes their positions. 
During this packing phase, we simply pseudorandomly map \texttt{design} \texttt{SiteInst}s onto \texttt{device} \texttt{Site}s in coordinate order as they are generated (e.g., the first \texttt{SiteInst} onto \texttt{SLICEL\_X0Y0}, the second onto \texttt{SLICEL\_X0Y1}, and so on).

One can think of a \texttt{design} \texttt{SiteInst} as a movable square peg and the corresponding \texttt{device} \texttt{Site} as a fixed square hole. 
Both \texttt{SiteInst}s and \texttt{Site}s come in various shapes or types (\texttt{SLICEL}, \texttt{SLICEM}, \texttt{DSP48E1}, etc.) and placement is only allowed between compatible pairs. 
For example, a \texttt{RAMB48E1} \texttt{SiteInst} can only be placed on a \texttt{RAMB48E1} \texttt{Site}.
A \texttt{DSP48E1} \texttt{SiteInst} can only be placed on a \texttt{DSP48E1} \texttt{Site}.
A \texttt{SLICEL} \texttt{SiteInst} may occupy either a \texttt{SLICEL} or a \texttt{SLICEM} \texttt{Site}, whereas a \texttt{SLICEM} \texttt{SiteInst} can only be placed on a \texttt{SLICEM} \texttt{Site}.
These constraints must be followed as the \texttt{SiteInst}s move across the device throughout the placement stage to ensure legality.

@@

Example:
    SiteInst si = new SiteInst("mySiteInst", design, SiteTypeEnum.SLICEL, device.getSite("SLICEL_X0Y1"));
    si.createCell(someFDRECell, si.getBEL("AFF"));
    si.routeSite();
    si.unplace();
    si.place(device.getSite("SLICEL_X15Y33"));
    // In Simulated Annealing, SiteInst objects will be unplaced() and placed() many times to converge to an optimal solution. 
    // All Cell-BEL mapping and intra-Site routing is preserved when a SiteInst is moved.

@@

\begin{lstlisting}[language=java, caption={Packing an individual \texttt{CarryCellGroup} into one \texttt{SLICEL} \texttt{SiteInst}s.}, label={lst:single_carry_chains}]
protected String[] FF_BELS = new String[] { "AFF", "BFF", "CFF", "DFF" };
protected String[] LUT6_BELS = new String[] { "A6LUT", "B6LUT", "C6LUT", "D6LUT" };

private void packCarrySite(CarryCellGroup carryCellGroup, SiteInst si) {
    // potential bug: what guarantees that all of the FFs connected to the CARRY4 all share the same CE and Reset?
    for (int i = 0; i < 4; i++) {
        EDIFHierCellInst ff = carryCellGroup.ffs().get(i);
        if (ff != null)
            si.createCell(ff, si.getBEL(FF_BELS[i]));
        EDIFHierCellInst lut = carryCellGroup.luts().get(i);
        if (lut != null)
            si.createCell(lut, si.getBEL(LUT6_BELS[i]));
        // carry site LUTs MUST be placed on LUT6 BELs.
        // only LUT6/O6 can connect to CARRY4/S0
    }
    si.createCell(carryCellGroup.carry(), si.getBEL("CARRY4"));
    // default intrasite routing
    si.routeSite();
    // sometimes the default routeSite() is insufficient, so some manual
    // intervention is required
    rerouteCarryNets(si);
    rerouteFFClkSrCeNets(si);
} // end placeCarrySite()
\end{lstlisting}

\newpage
\begin{lstlisting}[language=java, caption={Packing \texttt{CarryCellGroup}s into \texttt{SLICEL} \texttt{SiteInst}s.}, label={lst:carry_chains}]
private List<List<SiteInst>> packCarryChains(List<List<CarryCellGroup>> EDIFCarryChains)
        throws IOException {
    List<List<SiteInst>> siteInstChains = new ArrayList<>();
    writer.write("\n\nPacking carry chains... (" + EDIFCarryChains.size() + ")");
    for (List<CarryCellGroup> edifChain : EDIFCarryChains) {
        List<SiteInst> siteInstChain = new ArrayList<>();
        writer.write("\n\t\tChain Size: (" + edifChain.size() + "), Chain Anchor: "
                + edifChain.get(0).carry().getFullHierarchicalInstName());
        Site anchorSite = selectCarryAnchorSite(edifChain.size());
        SiteTypeEnum selectedSiteType = anchorSite.getSiteTypeEnum();
        for (int i = 0; i < edifChain.size(); i++) {
            Site site = (i == 0) ? anchorSite
                    : device.getSite("SLICE_X" + anchorSite.getInstanceX() + "Y" + (anchorSite.getInstanceY() + i));
            SiteInst si = new SiteInst(edifChain.get(i).carry().getFullHierarchicalInstName(), design,
                    selectedSiteType,
                    site);
            packCarrySite(edifChain.get(i), si);
            if (i == 0) { // additional routing logic for anchor site
                Net CINNet = si.getNetFromSiteWire("CIN");
                CINNet.removePin(si.getSitePinInst("CIN"));
                si.addSitePIP(si.getSitePIP("PRECYINIT", "0"));
            }
            occupiedSites.get(selectedSiteType).add(site);
            availableSites.get(selectedSiteType).remove(site);
            siteInstChain.add(si);
        }
        siteInstChains.add(siteInstChain);
    } // end for (List<EDIFCellInst> chain : EDIFCarryChains)
    return siteInstChains;
} // end packCarryChains()

\begin{lstlisting}[language=java, caption={Manually rerouting intra-Site nets in a \texttt{SLICEL} containing \texttt{CARRY4}}]
    protected String[] FF_BELS = new String[] { "AFF", "BFF", "CFF", "DFF" };
    private void rerouteCarryNets(SiteInst si) {
        // activate PIPs for CARRY4/COUT
        si.addSitePIP(si.getSitePIP("COUTUSED", "0"));
        // undo default CARRY4/DI nets
        SitePinInst AX = si.getSitePinInst("AX");
        if (AX != null)
            si.unrouteIntraSiteNet(AX.getBELPin(), si.getBELPin("ACY0", "AX"));
        SitePinInst DX = si.getSitePinInst("DX");
        if (DX != null)
            si.unrouteIntraSiteNet(DX.getBELPin(), si.getBELPin("DCY0", "DX"));
        // activate PIPs for CARRY4/DI pins
        si.addSitePIP(si.getSitePIP("DCY0", "DX"));
        si.addSitePIP(si.getSitePIP("CCY0", "CX"));
        si.addSitePIP(si.getSitePIP("BCY0", "BX"));
        si.addSitePIP(si.getSitePIP("ACY0", "AX"));
        // remove stray CARRY4/CO nets
        if (si.getNetFromSiteWire("CARRY4_CO2") != null)
            design.removeNet(si.getNetFromSiteWire("CARRY4_CO2"));
        if (si.getNetFromSiteWire("CARRY4_CO1") != null)
            design.removeNet(si.getNetFromSiteWire("CARRY4_CO1"));
        if (si.getNetFromSiteWire("CARRY4_CO0") != null)
            design.removeNet(si.getNetFromSiteWire("CARRY4_CO0"));
        // add default XOR PIPs for unused FFs
        for (String FF : FF_BELS)
            if (si.getCell(FF) == null)
                si.addSitePIP(si.getSitePIP(FF.charAt(0) + "OUTMUX", "XOR"));
    } // end rerouteCarryNets()
\end{lstlisting}
\begin{multicols}{2}

\subsection{Placement}
    \label{subsec:placement}
    Up until now we have only organized the logical \texttt{EDIFHierCellInst}s into \texttt{SiteInst}s. 
    This is where simulated annealing actually begins where we actually place the \texttt{SiteInst}s onto physical \texttt{Site}s on the \texttt{device} level.
